---
title: "Everything is forcats"
subtitle: "Working with factors"
author: "Zaira Gonzalez Mora & Luka Vasilj"
output: 
  html_document:
    toc: TRUE
    df_print: paged
    number_sections: FALSE
    highlight: tango
    theme: lumen
    toc_depth: 3
    toc_float: true
    css: custom.css 
    self_contained: false
    includes:
      after_body: footer.html
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction

What will be covered in the workshop:
- what is categorical data?
- What are factors?
- Why are they useful?
- How do we incorporate factors into the concept of tidy data?

```{r}

```

## Categorical Data

**What is a Categorical Variable?**
<br>
- Definition: Variables that have a fixed and known set of possible values.[source: https://forcats.tidyverse.org/]

- Issue: categorical variables are often coded with numerical values.

- The analysis of categorical data generally involves the use of tables by counting the number of observations that fall into categories.

- For example race, age, group, educational level, income brackets. [Source: http://www.stat.yale.edu/Courses/1997-98/101/catdat.htm]
<br>

## Factors
<br>
- R uses factors to handle categorical variables.
- R represents categorical variables with factors.
- Factors could be easier to work with than characters.
- A Factor is an integer vector with a levels attribute that stores a set of mappings between integers and categorical values.
- Factors are helpful for reordering character vectors in non alphabetical order and to improve display.
- Categorical data is often coded with numerical values.

## Application
- Wrangling of categorical data can be difficult.
- "Careless data cleaning operations can lead to errors or inconsistencies in analysis" (McNamara et all, 2007)


## Forcats
- It is better to work with categorical data using tidyverse than with base R. 
- Why? Tidyverse is more human readable.
- HOWEVER, forcats is not part of the core tidyverse as we need to load it explicitly.
- When you view a factor, R does not display the integers but the values associated with them.
- The Forecats package provides tools for working with factors, which (again) are R's data structure for categorical data.
- 

```{r pressure, echo=FALSE}
#Potentially include a GIF that shows forcats is an anagram of factor.

### I couldn't find something funny. BUT, we can say this:

# Forecats: *For* working with *cat*egorical variable*s*

#We could create something like this (just for you to see the little image on the section How to solve an anagram in 6 steps): https://anagramscramble.com/anagrams-of/gifts

#I am not sure if we have the time but we could try to change one of the gifs here with our word (READ/DEAR image for instance or Ladies Rest/Lead sister or CAB DRIVER/CRAB DIVER): https://giphy.com/search/anagram
### I remember changing legends of memes in stats II but I never did it on a gif.

#This one could also work: https://giphy.com/gifs/season-16-the-simpsons-16x16-xT5LMIPOBXmmVuoLQI 

```


